\section{Conclusion}

This project presented \textbf{FitCoachAR}, a real-time, adaptive exercise coaching system that bridges the gap between academic pose-analysis pipelines and practical consumer fitness applications.

Through a comprehensive empirical evaluation, we showed that among a wide range of pose estimation backends,
\textbf{MoveNet\_2D} offers the most favorable accuracy--latency trade-off for real-time use.
More importantly, our four-step post-processing pipeline—body-centric canonicalization,
yaw compensation, confidence-/motion-adaptive temporal smoothing, and anatomical consistency projection—
substantially improves both joint-angle accuracy and cross-view stability without increasing inference latency.

Beyond low-level pose accuracy, FitCoachAR introduces \textbf{FormScript}, a semantic intermediate representation that transforms continuous pose measurements into interpretable FormCodes.
This abstraction enables exercise-specific reasoning, personalized calibration, and natural-language feedback,
supporting both real-time corrective cues and post-workout summaries powered by \textbf{LLM-based coaching} (Llama 3.3 70B via Cerebras).
We validated the FormCode system on the RepCount dataset (101 squat videos, 969 repetitions), achieving biomechanically plausible classifications through data-driven threshold optimization.
By decoupling biomechanical definitions from pose estimation and feedback generation,
FormScript provides a flexible and extensible framework for future exercise analysis systems.

Overall, FitCoachAR illustrates a practical design philosophy for interactive human-motion applications:
prioritizing robustness, interpretability, and responsiveness over raw model complexity.
While several limitations remain—including the lack of ablation analysis and the reliance on aggregated metrics—
the results suggest that integrated, lightweight post-processing can significantly narrow the gap
between small 2D models and heavier 3D approaches in real-world fitness coaching scenarios.

Future work may explore finer-grained phase-specific evaluation, user studies on coaching effectiveness,
and the extension of FormCodes to rehabilitation and clinical movement assessment.
For FormScript specifically, future directions include developing personalized calibration for all FormCodes (not just range-of-motion metrics), expanding the exercise library beyond squats and bicep curls, and refining the LLM prompting strategy to generate more targeted coaching recommendations.
Additionally, evolutionary algorithm (EA) approaches could automate threshold calibration without hardcoding exercise-specific parameters, and classification methods could be employed to automatically identify which joints are most biomechanically significant for each exercise type.
Nonetheless, this project demonstrates that real-time, personalized, and explainable exercise coaching
is achievable on consumer hardware using principled pose processing and semantic reasoning.
