\section{Division of Labor}
\label{sec:division-of-labor}

This project was completed as a group effort. Below we summarize the workload division.

\paragraph{Yangyang Zhang.}
Yangyang's primary responsibilities include:
\begin{itemize}
    \item Establishing the core frontend--backend streaming pipeline (e.g., WebSocket-based frame delivery and latency reduction), along with primary front-end UI design and the initial AR overlay/markers.
    \item Design and proposal of the four lightweight post-processing techniques (reported in Sec.~\ref{sec:model-eval}) for stabilizing MoveNet~2D outputs.
    \item Empirical model analysis and experiments reported in Sec.~\ref{sec:movenet-postproc}, covering multi-backend benchmarking, stability evaluation, and latency measurement.
\end{itemize}

\paragraph{Seyedmohamad Mirhoseininejad.}
Seyedmohamad's primary responsibilities include:
\begin{itemize}
    \item Design and implementation of the FormScript feedback system, including elementary FormCodes and Super FormCodes for rule-based form analysis.
    \item Integration of LLM-powered post-workout summaries using Llama 3.3 70B via Cerebras, including per-rep feedback aggregation.
    \item FormCode validation experiments on the RepCount dataset (Sec.~4.6).
\end{itemize}

\paragraph{Maximilian Fuchs.}
Maximilian Fuchs’s primary responsibilities include:
\begin{itemize}
    \item Design and implementation of the user-specific calibration process to derive exercise-dependent range-of-motion and timing parameters ($\theta_{\text{low}}$, $\theta_{\text{high}}$, $t_{\text{med}}$, $t_{\text{min}}$, $t_{\text{max}}$), along with auxiliary profiles (capabilities, tempo, symmetry, and form-constraint statistics), and persistent storage and activation of calibration records.
    \item Design and implementation of the normalized repetition counting module (\texttt{rep\_counter\_v2.py}), including an online state-machine for cycle detection driven by calibrated thresholds, normalized progress estimation, and timing-based robustness checks.
    \item Refactoring and integration of multiple 2D pose estimators into a unified, estimator-agnostic backend pipeline (\texttt{pose\_processor\_2d.py}), eliminating duplicated command and frame-processing logic and improving code maintainability without introducing new features.
\end{itemize}
